% spellchecker: disable

\documentclass[a4paper,twoside,10.5pt]{report} %openright

\input{Setup/Statics.tex}
\input{Setup/Preamble.tex}
\input{Setup/Settings.tex}

\begin{document}

\pagenumbering{roman}
\input{Frontmatter/Frontpage.tex}
\pagecolor{white}
\newgeometry{top=2.81cm, bottom=2.75cm, outer=2.5cm, inner=3.5cm}
\pagestyle{empty}
\cleardoublepage 
\input{Frontmatter/Copyright.tex}
\clearpage 
\pagestyle{main}
\input{Frontmatter/Approval.tex}
\clearpage 
\input{Frontmatter/Abstract.tex}
\clearpage 
\input{Frontmatter/Acknowledgements.tex}
\cleardoublepage 


\tableofcontents
\cleardoublepage 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagenumbering{arabic}
\include{Chapters/01_Introduction}
\include{Chapters/03_Inference}
\include{Chapters/04_Implementation}
\include{Chapters/05_Experiments}

\chapter{Conlusion}

\section{Conclusion}
We can get some(?) or not? improvement with respect to uncertianty estimation.

one stop shop for all mean

performance?

estimatino variance helps?


\section{Discussion}
Applicability of mcmc methods for non-deep learning with lots of data -> registerforskning fx.

what does it even mean to have a prior. correcntness of model

Reweightnig/resmapling of samplesaccording to validation.

how much does the correctness of the model matter -> model is obv. wrong...

various problems with hparam search approach
But prabably relavent for wanting to fit a model given a finitiate evalition budget
all seem to be overfitting <- fine in the sense that we dont want to continue training. However probably is biased towards qui

identificability huge search spaces

discuss implementation
 - Robust -- currently adhoc blablabla


ineteristing to train from scratch? smaller models worked fine -> batch norm... proably doesnt do muich for smaller models compared to deep models
-> dirty likelihoods


VI methods -> flows?
bunch of other stuff we could try to improve the performance of any of the methods lr scheduling, feature engineering, other model structures etc. 
dropout for prob method per dirty likelihood.

lack of modelling techniques?

%%%%%%%%%%%%%

\nocite{*}
\printbibliography[heading=bibintoc,title={Bibliography}]
\cleardoublepage 
\appendix
\input{Backmatter/Appendix.tex}
\cleartoleftpage
\input{Backmatter/Backpage.tex}

\end{document}